\documentclass[a4paper,landscape,10pt]{article}
\usepackage[top=1cm,left=1cm,right=1cm,bottom=2cm]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[lowtilde]{url}
\usepackage{color}
\usepackage[colorlinks]{hyperref}
\usepackage{scalerel}
\usepackage{multicol}
\usepackage[numbers,sort&compress]{natbib}
\bibliographystyle{plainnat}
\usepackage[swedish,finnish,french,spanish,latin,english]{babel}

\newcommand{\tblock}[5]{\noindent\begin{minipage}[t]{0.18\textwidth}\foreignlanguage{english}{\vspace*{1ex} #1\vspace*{1ex}}\end{minipage}\hskip 0.025\textwidth\begin{minipage}[t]{0.18\textwidth}\foreignlanguage{swedish}{\vspace*{1ex}#2\vspace*{1ex}}\end{minipage}\hskip 0.025\textwidth\begin{minipage}[t]{0.18\textwidth}\foreignlanguage{finnish}{\vspace*{1ex}#3\vspace*{1ex}}\end{minipage}\hskip 0.025\textwidth\begin{minipage}[t]{0.18\textwidth}\foreignlanguage{french}{\vspace*{1ex}#4\vspace*{1ex}}\end{minipage}\hskip 0.025\textwidth\begin{minipage}[t]{0.18\textwidth}\foreignlanguage{spanish}{\vspace*{1ex}#5\vspace*{1ex}}\end{minipage}}

%\addtolength{\oddsidemargin}{-1in}
%\addtolength{\evensidemargin}{-1in}
%\addtolength{\textwidth}{2in}

%\usepackage[makeindex]{glossaries}
%\usepackage[automake]{glossaries-extra}
%\makeglossaries
%\loadglsentries{bor-glossary}


%opening
\title{\huge \textbf{U.~S.\ Declaration of Independence}}
\author{---}
\date{July 4, 1776}
\begin{document}
	
\sloppy
\maketitle
\begin{abstract}
	\centering
	English \cite{gov1}, with translations to Swedish, Finnish, French, \& Spanish.
	
	\copyright\ Public domain: all translations.
	
\end{abstract}

\tblock{In Congress, July 4, 1776}
{I Samlag, 4 juli 1776}
{Eduskunnassa, neljännellä maaliskuuta, vuonna 1776}
{En Congrès, 4 juli 1776}
{En Congreso, 4 juli 1776}
\tblock{The unanimous Declaration of the thirteen united States of America,}
{Den enige Förklaringen av de tretton Amerikas Förenta Stater,}
{Yksimielinen julistus kolmestatoista Amerikan yhdistetystä vallasta,}
{---}
{---}
{---}


\tblock{When in the Course of human events,}
{När i gången av mänskliga händelser,}
{Koska inhimillisten tapahtumisten käymisessä,}
{Lorsqu'en train des évennements humains,}
{---}

\tblock{it becomes necessary for one people to dissolve the political bands which have connected them with another,}
{det bliver nödvändigt av ett folk att skilja de politiska band, vilka hava förbindat dem med ett annat,}
{tulee tarpeelliseksi yhden kansan purkaa niitä poliittisia siteitä, jotka ovat sitoneet heidän toiseen,}
{---}{---}


\tblock{and to assume among the powers of the earth,}
{och att antaga ibland jordens makter,}
{ja ottaa vastaan maan valtien seassa,}
{---}{---}

\tblock{the separate and equal station to which the Laws of Nature and of Nature's God entitle them,}
{den särskilda och lika stationen till vilken Naturens och Naturens Guds Lag berättiga dem,}
{erillistä ja kaltaista sijaa johon luonnon laki ja luonnon Jumalan laki oikeuttavat heidän,}{---}{---}

\tblock{a decent respect to the opinions of mankind}
{---}{---}{---}{---}


\tblock{requires that they should declare the causes which impel them to the separation.}
{---}{---}{---}{---}

\tblock{We hold these truths to be self-evident,}
{Vi hålla desse sanningar att vara själv-klare,}
{Pidetään näitä totuuksia itsestään selvinä,}{---}{---}
	
\tblock{that all men are created equal,}
{att alla människor äro skapade lika,}
{että kaikki ihmiset on luotu kaltaisiksi,}{---}{---}

\tblock{that they are endowed by their Creator with certain unalienable Rights,}
{att de äro förlänade av sin Skapare med visse oförytterlige rättigheter,}
{että heidän on annettu Luojaltaan eräitä siirtämättömiä oikeuksia,}{---}{---}

\tblock{that among these are Life, Liberty and the pursuit of Happiness.}
{att ibland desse äro Liv, Frihet, och Lyckans eftersträvande}{että näiden seassa on elämä, vapaus, ja pyrkiminen Onnellisuuteen.}{---}{---}

\tblock{---That to secure these rights, Governments are instituted among Men,}
{---På det för att försäkra desse rättigheter äro Regeringar stiftade ibland Människor,}
{---Että varmistaa näitä oikeuksia, on Hallituksia perustettu ihmisten seassa,}{---}{---}

\tblock{deriving their just powers from the consent of the governed,}
{åstadkommande sine rättvise makter från de Regerades myndighet,}
{johtuen oikeita valtuuksia Hallittujen hyväksymyksestä}{---}{---}

\tblock{---That whenever any Form of Government becomes destructive of these ends,}
{---På det att när som helst någon Form av Regering bliver förstörande av dessa ändamål,}
{---Että koska ikänä mitään Hallituksen muotoa tulee hävittämään näitä syitä,}{---}{---}

\tblock{it is the Right of the People to alter or to abolish it,}
{så är det Folkets Rätt att förändra eller avskaffa den,}
{---}{---}{---}

\tblock{and to institute new Government,}
{och at inrätta ny Regering,}{---}{---}{---}

\tblock{laying its foundation on such principles and organizing its powers in such form,}
{laggande dess grund på sådane principer och ordnande dess makter i sådan form,}{}{}{}


\tblock{as to them shall seem most likely to effect their Safety and Happiness.}
{som till dem skall tyckas sannolikast åstadkomma sin Trygghet och Lycka.}{}{}{}

\tblock{Prudence, indeed, will dictate that Governments long established should not be changed for light and transient causes;}
{Forsiktighet, i sanningen, vill beordra att Regeringar länge inrättade icke skulle ändras fór lätta ock flyktiga mål}{---}{---}{---}

\tblock{and accordingly all experience hath shewn,}
{och i enighet därmed har all erfarenhet bevisat,}
{---}{---}{---}

\tblock{that mankind are more disposed to suffer,}
{att människor äro vanare att lida,}
{ }{---}{---}

\tblock{while evils are sufferable,}
{så länge onda ting äro liderliga,}{---}{---}{---}

\tblock{than to right themselves by abolishing the forms to which they are accustomed.}
{än att rätta sig genom att avskaffa de former vid vilka de hava blivit vana}{---}{---}{---}

\tblock{But when a long train of abuses and usurpations,}
{Men då ett långt tåg missbruk ock tillskansande,}{---}{---}{---}

\tblock{pursuing invariably the same Object}
{eftersträvande allesammans samma målet}{---}{---}{---}

\tblock{evinces a design to reduce them under absolute Despotism,}
{ådagalaggar en komplott att underkasta dem till absolut despotism,}{---}{---}{---}

\tblock{it is their right,}
{så är det deras rätt,}{---}{---}{---}

\tblock{it is their duty,}{så är det deras plikt,}{---}{---}{---}
\tblock{to throw off such Government,}{att avkasta sådan Regering,}{---}{---}{---}
\tblock{and to provide new Guards for their future security.}{och att förse nye Åtgärder till sin framtide säkerhet.}{---}{---}{---}
\tblock{---Such has been the patient sufferance of these Colonies;}{Sådan har varit desse koloniers långmodige lidelse;}{---}{---}{---}

\tblock{and such is now the necessity which constrains them to alter their former Systems of Government.}{och sådan är nu nödvändigheten som tvinger dem att ändra sina tidigare System av Regering.}{---}{---}{---}
\tblock{The history of the present King of Great Britain is a history of repeated injuries and usurpations,}{Historien av den nuvarande Kungen av Stor Brittanien är en historia av upprepade skador och tillskansande,}{---}{---}{---}
\tblock{all having in direct object the establishment of an absolute Tyranny over these States.}{allesammans havande i direkt syfte inrättelsen av ett absolut tyranni över desse Stater.}{---}{---}{---}
\tblock{To prove this, let Facts be submitted to a candid world.}{---}{---}{---}{---}

\tblock{He has refused his Assent to Laws,}{---}{---}{---}{---}
\tblock{the most wholesome and necessary for the public good.}{---}{---}{---}{---}

\tblock{He has forbidden his Governors to pass Laws of immediate and pressing importance,}{---}{---}{---}{---}
\tblock{unless suspended in their operation till his Assent should be obtained;}{---}{---}{---}{---}
\tblock{and when so suspended, he has utterly neglected to attend to them.}{---}{---}{---}{---}

\tblock{He has refused to pass other Laws for the accommodation of large districts of people,}{---}{---}{---}{---}
\tblock{unless those people would relinquish the right of Representation in the Legislature,}{---}{---}{---}{---}
\tblock{a right inestimable to them and formidable to tyrants only.}{---}{---}{---}{---}

\tblock{He has called together legislative bodies at places unusual, uncomfortable, and distant from the depository of their public Records,}{---}{---}{---}{---}
\tblock{for the sole purpose of fatiguing them into compliance with his measures.}{---}{---}{---}{---}

\tblock{He has dissolved Representative Houses repeatedly,}{---}{---}{---}{---}
\tblock{for opposing with manly firmness his invasions on the rights of the people.}{---}{---}{---}{---}

\tblock{He has refused for a long time, after such dissolutions, }{---}{---}{---}{---}
\tblock{to cause others to be elected;}{---}{---}{---}{---}
\tblock{whereby the Legislative powers, incapable of Annihilation,}{---}{---}{---}{---}
\tblock{have returned to the People at large for their exercise;}{---}{---}{---}{---}
\tblock{the State remaining in the mean time exposed to all the dangers}{---}{---}{---}{---}
\tblock{of invasion from without, and convulsions within.}{---}{---}{---}{---}

\tblock{He has endeavoured to prevent the population of these States;}{---}{---}{---}{---}
\tblock{for that purpose obstructing the Laws for Naturalization of Foreigners;}{---}{---}{---}{---}
\tblock{refusing to pass others to encourage their migrations hither,}{---}{---}{---}{---}
\tblock{and raising the conditions of new Appropriations of Lands.}{---}{---}{---}{---}

\tblock{He has obstructed the Administration of Justice,}{---}{---}{---}{---}
\tblock{by refusing his Assent to Laws for establishing Judiciary powers.}{---}{---}{---}{---}

\tblock{He has made Judges dependent on his Will alone,}{---}{---}{---}{---}
\tblock{for the tenure of their offices,}{---}{---}{---}{---}
\tblock{and the amount and payment of their salaries.}{---}{---}{---}{---}

\tblock{He has erected a multitude of New Offices,}{---}{---}{---}{---}
\tblock{and sent hither swarms of Officers to harrass our people,}{---}{---}{---}{---}
\tblock{and eat out their substance.}{---}{---}{---}{---}

\tblock{He has kept among us, in times of peace,}{---}{---}{---}{---}
\tblock{Standing Armies without the Consent of our legislatures.}{---}{---}{---}{---}

\tblock{He has affected to render the Military independent of and superior to the Civil power.}{---}{---}{---}{---}

\tblock{He has combined with others to subject us to a jurisdiction foreign to our constitution,}{---}{---}{---}{---}
\tblock{and unacknowledged by our laws; giving his Assent to their Acts of pretended Legislation:}{---}{---}{---}{---}

\tblock{For Quartering large bodies of armed troops among us:}{---}{---}{---}{---}

\tblock{For protecting them, by a mock Trial,}{---}{---}{---}{---}
\tblock{from punishment for any Murders}{---}{---}{---}{---}
\tblock{which they should commit on the Inhabitants of these States:}{---}{---}{---}{---}

\tblock{For cutting off our Trade with all parts of the world:}{---}{---}{---}{---}

\tblock{For imposing Taxes on us without our Consent:}{---}{---}{---}{---}

\tblock{For depriving us in many cases, of the benefits of Trial by Jury:}{---}{---}{---}{---}

\tblock{For transporting us beyond Seas to be tried for pretended offences}{---}{---}{---}{---}

\tblock{For abolishing the free System of English Laws in a neighbouring Province,}{---}{---}{---}{---}
\tblock{establishing therein an Arbitrary government,}{---}{---}{---}{---}
\tblock{and enlarging its Boundaries so as to render it at once an example and fit instrument}{---}{---}{---}{---}
\tblock{for introducing the same absolute rule into these Colonies:}{---}{---}{---}{---}

\tblock{For taking away our Charters,}{---}{---}{---}{---}
\tblock{abolishing our most valuable Laws,}{---}{---}{---}{---}
\tblock{and altering fundamentally the Forms of our Governments:}{---}{---}{---}{---}

\tblock{For suspending our own Legislatures,}{---}{---}{---}{---}
\tblock{and declaring themselves invested with power to legislate for us in all cases whatsoever.}{---}{---}{---}{---}

\tblock{He has abdicated Government here,}{---}{---}{---}{---}
\tblock{by declaring us out of his Protection and waging War against us.}{---}{---}{---}{---}

\tblock{He has plundered our seas,}{---}{---}{---}{---}
\tblock{ravaged our Coasts,}{---}{---}{---}{---}
\tblock{burnt our towns,}{---}{---}{---}{---}
\tblock{and destroyed the lives of our people.}{---}{---}{---}{---}

\tblock{He is at this time transporting large Armies of foreign Mercenaries}{---}{---}{---}{---}
\tblock{to compleat the works of death, desolation and tyranny,}{---}{---}{---}{---}
\tblock{already begun with circumstances of Cruelty \& perfidy}{---}{---}{---}{---}
\tblock{scarcely paralleled in the most barbarous ages,}{---}{---}{---}{---}
\tblock{and totally unworthy the Head of a civilized nation.}{---}{---}{---}{---}

\tblock{He has constrained our fellow Citizens taken Captive on the high Seas}{---}{---}{---}{---}
\tblock{to bear Arms against their Country,}{---}{---}{---}{---}
\tblock{to become the executioners of their friends and Brethren,}{---}{---}{---}{---}
\tblock{or to fall themselves by their Hands.}{---}{---}{---}{---}

\tblock{He has excited domestic insurrections amongst us,}{---}{---}{---}{---}
\tblock{and has endeavoured to bring on the inhabitants of our frontiers,}{---}{---}{---}{---}
\tblock{the merciless Indian Savages,}{---}{---}{---}{---}
\tblock{whose known rule of warfare,}{---}{---}{---}{---}
\tblock{is an undistinguished destruction of all ages, sexes and conditions.}{---}{---}{---}{---}

\tblock{In every stage of these Oppressions}{---}{---}{---}{---}
\tblock{We have Petitioned for Redress in the most humble terms:}{---}{---}{---}{---}
\tblock{Our repeated Petitions have been answered only by repeated injury.}{---}{---}{---}{---}
\tblock{A Prince whose character is thus marked by every act which may define a Tyrant,}{---}{---}{---}{---}
\tblock{is unfit to be the ruler of a free people.}{---}{---}{---}{---}

\tblock{Nor have We been wanting in attentions to our Brittish brethren.}{Icke heller hava vi saknat uppmärksamheter på våre Brittiske bröder.}{---}{---}{---}
\tblock{We have warned them from time to time}
{Vi hava varnat dem då och då}
{---}{---}{---}

\tblock{of attempts by their legislature to extend an unwarrantable jurisdiction over us.}
{om angrepp av deras lagstiftande församling att sträcka ut en obemyndigad domsrätt över oss.}{---}{---}{---}

\tblock{We have reminded them of the circumstances of our emigration and settlement here.}
{Vi hava påmint dem av omständigheterna kring vår utvandring och bosättning hit.}{---}{---}{---}

\tblock{We have appealed to their native justice and magnanimity,}
{Vi hava vädjat till deras inhemska rättvisa och storsinthet,}{---}{---}{---}
\tblock{and we have conjured them by the ties of our common kindred to disavow these usurpations,}{---}{---}{---}{---}
\tblock{which, would inevitably interrupt our connections and correspondence.}{---}{---}{---}{---}
\tblock{They too have been deaf to the voice of justice and of consanguinity.}
{De också hava varit döve till rättvisans och blodsfrändskapets röst.}{---}{---}{---}
\tblock{We must, therefore, acquiesce in the necessity,}{Vi måste, därför, tilltyga i nödvändigheten}{---}{---}{---}
\tblock{which denounces our Separation,}{som fördömer vår skilsmässa,}{---}{---}{---}
\tblock{and hold them, as we hold the rest of mankind,}{och hålla dem, som vi hålla det övriga av människosläktet,}{---}{---}{---}
\tblock{Enemies in War, in Peace Friends.}{ovänner i krig, i fred vänner.}{---}{---}{---}

\tblock{We, therefore, the Representatives of the united States of America,}
{Vi, därför, Amerikas Förenta Staternas Representater,}
{Sentähden, Edustajailta Amerikan Yhdistetuistä Valtoista,}{---}{---}
\tblock{in General Congress, Assembled,}{i allmänt Samlag, församlade,}{yleisessä Eduskunnassa, kokoonnuttu}{en général Congrès, assemblées,}{---}
\tblock{appealing to the Supreme Judge of the world for the rectitude of our intentions,}
{---}
{vedottaessa maailman Korkeimpaan Tuomariin aikeiden vanhurskaudesta}{---}{---}
\tblock{do, in the Name, and by Authority of the good People of these Colonies, solemnly publish and declare,}{---}{näiden siirtomaiden hyvän kansan nimessä ja hyväksymisessä, vakavasti julkaistaan ja ilmoitetaan,}{---}{---}
\tblock{That these United Colonies are, and of Right ought to be Free and Independent States;}
{Att desse Förenta Colonier äro, och av rätt borde vara Frie och Oberoende Stater;}
{Että nämä siirtomaat ovat, ja oikeasti pitäisi heidän oleman, Vapaat ja Riippumattomat Vallat;}
{---}{---}
\tblock{that they are Absolved from all Allegiance to the British Crown,}{---}{---}{---}{---}
\tblock{and that all political connection between them and the State of Great Britain,}
{att all politisk förbindelse emellan dem och Stor-Brittaniens Stat,}{---}{---}{---}
\tblock{is and ought to be totally dissolved;}
{är och borde vara helt och hållet skild;}{---}{---}{---}
\tblock{and that as Free and Independent States,}
{och som Frie och Oberoende Stater,}{---}{---}{---}
\tblock{they have full Power}{så har de full Makt}{---}{---}{---}
\tblock{to levy War,}{att föra krig,}{---}{---}{---}
\tblock{conclude Peace,}{föredraga fred,}{---}{---}{---}
\tblock{contract Alliances,}{avtala fördrag,}{---}{---}{---}
\tblock{establish Commerce,}{inrätta handel,}{---}{---}{---}
\tblock{and to do all other Acts and Things which Independent States may of right do.}
{och föra alle andre Gärningar och Ting som Oberoende Stater må av rätt göra.}{---}{---}{---}
\tblock{And for the support of this Declaration,}{---}{---}{---}{---}
\tblock{with a firm reliance on the protection of divine Providence,}{---}{---}{---}{---}
\tblock{we mutually pledge to each other our Lives, our Fortunes and our sacred Honor.}{---}{---}{---}{---}
\pagebreak
\hrule
\begin{multicols}{7}
	\fussy
\subsection*{Georgia}

Button Gwinnett\\
Lyman Hall\\
George Walton

\subsection*{North Carolina}

William Hooper\\
Joseph Hewes\\
John Penn

\subsection*{South Carolina}

Edward Rutledge\\
Thomas Heyward, Jr.\\
Thomas Lynch, Jr.\\
Arthur Middleton

\subsection*{Massachusetts}

John Hancock

\subsection*{Maryland}

Samuel Chase\\
William Paca\\
Thomas Stone\\
Charles Carroll of Carrollton

\subsection*{Virginia}

George Wythe\\
Richard Henry Lee\\
Thomas Jefferson\\
Benjamin Harrison\\
Thomas Nelson, Jr.\\
Francis Lightfoot Lee\\
Carter Braxton

\subsection*{Pennsylvania}

Robert Morris\\
Benjamin Rush\\
Benjamin Franklin\\
John Morton\\
George Clymer\\
James Smith\\
George Taylor\\
James Wilson\\
George Ross

\subsection*{Delaware}

Caesar Rodney\\
George Read\\
Thomas McKean

\subsection*{New York}

William Floyd\\
Philip Livingston\\
Francis Lewis\\
Lewis Morris

\subsection*{New Jersey}

Richard Stockton\\
John Witherspoon\\
Francis Hopkinson\\
John Hart\\
Abraham Clark

\subsection*{New Hampshire}

Josiah Bartlett\\
William Whipple\\

\subsection*{Massachusetts}

Samuel Adams\\
John Adams\\
Robert Treat Paine\\
Elbridge Gerry


\subsection*{Rhode Island}

Stephen Hopkins\\
William Ellery



\subsection*{Connecticut}

Roger Sherman\\
Samuel Huntington\\
William Williams\\
Oliver Wolcott

\subsection*{New Hampshire}

Matthew Thornton
\end{multicols}
\nopagebreak
\hrule

%\printglossaries
\bibliography{doi}{}

\end{document}
